---
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---
```{r intro, echo=FALSE, results="hide"}
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE, 
                      comment = "", 
                      warning=FALSE, 
                      results="hide") 
knitr::opts_knit$set(root.dir = "C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_US/")

date<-as.Date(Sys.time(	), format='%d%b%Y')
time<-Sys.time()

suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(plotly)))
suppressWarnings(suppressMessages(library(Matrix)))
suppressWarnings(suppressMessages(library(stringr)))
suppressWarnings(suppressMessages(library(stringi)))

suppressWarnings(suppressMessages(library(jsonlite)))
suppressWarnings(suppressMessages(library(httr)))
suppressWarnings(suppressMessages(library(rlist)))
suppressWarnings(suppressMessages(library(zoo)))
suppressWarnings(suppressMessages(library(RColorBrewer))) 
suppressWarnings(suppressMessages(library(lubridate)))
```

```{r dataPop}
# https://www.census.gov/newsroom/press-kits/2020/population-estimates-detailed.html
# Annual Estimates of the Resident Population by Sex, Age, Race, and Hispanic Origin for the United States: April 1, 2010 to July 1, 2019
# NC-EST2019-ASR6H

dtapop<-read.csv("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_US/DataCensusBureau/nc-est2019-asr6h-2019.csv")
names(dtapop)<- tolower(names(dtapop))
    colnames(dtapop) 
    table(dtapop$race) 
    table(dtapop$hispanicorigin) 
 
    #View(dtapop)

dtapopage<-dtapop%>%
    filter(sex=="Total")%>%
    filter(race=="Total")%>%
    filter(hispanicorigin=="Total")%>%
    select(race, hispanicorigin, sex, starts_with("age"), pop)%>%
    filter(agegroup_5=="yes")

    sum(dtapopage$pop)

dtapopage10<-dtapopage%>%
    mutate(
        age=ifelse(age== 5,  0, age),
        age=ifelse(age==15, 10, age),
        age=ifelse(age==25, 20, age),
        age=ifelse(age==35, 30, age),
        age=ifelse(age==45, 40, age),
        age=ifelse(age==55, 50, age),
        age=ifelse(age==65, 60, age),
        age=ifelse(age==75, 70, age),
        age=ifelse(age==85, 80, age)
    )%>%
    group_by(race, hispanicorigin, sex, age)%>%
    summarise_at(vars(pop), funs(sum))
    
    sum(dtapopage10$pop)

dtapopagerace<-dtapop%>%
    filter(sex=="Total")%>%
    filter(race!="Total")%>%
    filter(hispanicorigin!="Total")%>%
    select(race, hispanicorigin, sex, starts_with("age"), pop)%>%
    filter(agegroup_5=="yes")%>%
    mutate(
        racegroup="", 
        racegroup=ifelse((race=="White" & hispanicorigin=="Not Hispanic"),
                         "NH White", racegroup), 
        racegroup=ifelse((race=="Black" & hispanicorigin=="Not Hispanic"),
                         "NH Black", racegroup), 
        racegroup=ifelse((race=="Asian" & hispanicorigin=="Not Hispanic"),
                         "NH Asian", racegroup), 
        racegroup=ifelse((race=="AIAN" & hispanicorigin=="Not Hispanic"),
                         "NH AIAN", racegroup), 
        racegroup=ifelse((race=="NHPI" & hispanicorigin=="Not Hispanic"),
                         "NH NHPI", racegroup), 
        racegroup=ifelse((race=="TwoOrMoreRaces" & hispanicorigin=="Not Hispanic"),
                         "NH Multiple/Other", racegroup), 
        racegroup=ifelse((hispanicorigin=="Hispanic"),
                         "Hispanic", racegroup)
    )%>%
    select(racegroup, starts_with("age"), pop)%>%
    group_by(racegroup, age, agegroup)%>%
    summarise_at(vars(pop), funs(sum))%>%
    ungroup()

    sum(dtapopagerace$pop)
    
dtapopagerace10<-dtapopagerace%>%
    mutate(
        age=ifelse(age== 5,  0, age),
        age=ifelse(age==15, 10, age),
        age=ifelse(age==25, 20, age),
        age=ifelse(age==35, 30, age),
        age=ifelse(age==45, 40, age),
        age=ifelse(age==55, 50, age),
        age=ifelse(age==65, 60, age),
        age=ifelse(age==75, 70, age),
        age=ifelse(age==85, 80, age)
    )%>%
    group_by(racegroup, age)%>%
    summarise_at(vars(pop), funs(sum))    

    sum(dtapopagerace10$pop)

```

```{r CaseSurveillanceAPI, eval=FALSE}
# COVID-19 Case Surveillance Public Use (case level)
# https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf
    
    ##### API 
    
url<-("https://data.cdc.gov/resource/vbim-akqf.json")

jsondata<-fromJSON(url) 
    str(jsondata)
    
dta<-jsondata

    colnames(dta)    
    
    table(dta$current_status)
    table(dta$age_group)
    table(dta$sex)
    table(dta$race_ethnicity_combined)
    #table(dta$onset_dt)  
```

```{r CaseSurveillanceExport}

# COVID-19 Case Surveillance Public Use (case level)
# https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf

    ##### Export to CSV
    
dtasurveillance<-read.csv("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_US/DataCDC/COVID-19_Case_Surveillance_Public_Use_Data.csv")    
    names(dtasurveillance)<- tolower(names(dtasurveillance))
    colnames(dtasurveillance)
    nrow(dtasurveillance)

    #table(dtasurveillance$current_status)
    #table(dtasurveillance$age_group)
    #table(dtasurveillance$sex)
    #table(dtasurveillance$race.and.ethnicity..combined.)
    #table(dta$onset_dt)  
    #table(dta$pos_spec_dt)
    #table(dta$cdc_report_dt)    
    
# OMG age categories different for the case surveillance and tabulated death count data! 
# If we want to use same categoreis for cases and deaths, we must use the surveillance data. Ugh...
# Get pop data readu accordingly 10- year interval
```

```{r dtacase}
# CASE LEVEL data managment         
dtacase<-dtasurveillance%>%
    mutate_if(is.factor, as.character) %>% 
    mutate(
        
        cdc_report_dt=as.Date(cdc_report_dt),
        week=as.numeric(
            strftime(as.POSIXct(cdc_report_dt), format = "%V")),
        month=as.numeric(
            substr(as.character(as.Date(cdc_report_dt, format = "%m")), 6, 7)), 
        
        sextotal="Total",
        
        agegroup=age_group, 
            agegroup=ifelse(is.na(agegroup)==TRUE, "Unknown", agegroup),
        age=as.numeric(sapply(strsplit(agegroup,"-"), `[`, 1)), #age interval begins
            age=ifelse(agegroup=="80+ Years", 80, age),
        
        racegroup=race.and.ethnicity..combined., 
        racegroup=ifelse(race.and.ethnicity..combined.=="White, Non-Hispanic",
                         "NH White", racegroup), 
        racegroup=ifelse(race.and.ethnicity..combined.=="Black, Non-Hispanic",
                         "NH Black", racegroup), 
        racegroup=ifelse(race.and.ethnicity..combined.=="Asian, Non-Hispanic",
                         "NH Asian", racegroup), 
        racegroup=ifelse(race.and.ethnicity..combined.=="American Indian/Alaska Native, Non-Hispanic",
                         "NH AIAN", racegroup), 
        racegroup=ifelse(race.and.ethnicity..combined.=="Native Hawaiian/Other Pacific Islander, Non-Hispanic",
                         "NH NHPI", racegroup), 
        racegroup=ifelse(race.and.ethnicity..combined.=="Multiple/Other, Non-Hispanic",
                         "NH Multiple/Other", racegroup), 
        racegroup=ifelse(race.and.ethnicity..combined.=="Hispanic/Latino",
                         "Hispanic", racegroup),
        racegroup=ifelse(is.na(race.and.ethnicity..combined.)==TRUE,
                         "Unknown", racegroup),
        
        raceagegroup=paste0(racegroup, "_",agegroup), 

        missingage  =as.numeric(is.na(age)==TRUE),
        missingsex  =as.numeric(sex!="Male" & sex!="Female" ),
        missingrace =as.numeric(racegroup=="Unknown"),
        missingdeath=as.numeric(death_yn!="Yes" & death_yn!="No"),  
        missinghosp =as.numeric(hosp_yn!="Yes" & hosp_yn!="No"),
        
        case=1, 
        death=as.numeric(death_yn=="Yes"), 
        hosp =as.numeric(hosp_yn=="Yes"), 
        
            death=ifelse(missingdeath==1, NA, death), 
            hosp =ifelse(missinghosp ==1, NA, hosp)
        
        )

colnames(dtacase)    
str(dtacase)
#summary(dtacase)

#table(dtacase$raceagegroup)


```

```{r dtamonthly}
# Weekly trend count data by RACE AND AGE
dtamonthly<-dtacase%>%
    arrange(racegroup, agegroup, month)%>%
    mutate(raceagegroupmonth=paste0(racegroup,"_",agegroup,"_",as.character(month)))%>%
    group_by(raceagegroupmonth)%>%
    summarise_at(vars(case, death, hosp, missingdeath, missinghosp), funs(sum), na.rm = TRUE)%>%
    ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupmonth,"_"), `[`, 1),
        agegroup=sapply(strsplit(raceagegroupmonth,"_"), `[`, 2),
        month=as.numeric(sapply(strsplit(raceagegroupmonth,"_"), `[`, 3)), 
        
        age=as.numeric(sapply(strsplit(agegroup,"-"), `[`, 1)), #age interval begins
            age=ifelse(agegroup=="80+ Years", 80, age), 
        
        raceagegroup=paste0(racegroup, "_", agegroup) 
        )%>%
    arrange(racegroup, agegroup, month)
    

    #View(dtamonthly)
    #nrow(dtamonthly)
    #str(dtamonthly)
    #summary(dtamonthly)

dim(dtamonthly)
dtamonthly<-left_join(dtamonthly, dtapopagerace10, by = c("racegroup", "age")) 
dim(dtamonthly)    

write.csv(dtamonthly, "X_COVID19_US_AgeComposition/dtamonthly_20201204.csv")
```

```{r dtamonthlybyrace}
# Weekly trend count data by RACE AND AGE
dtamonthlybyrace<-dtamonthly%>%
    arrange(racegroup, agegroup, month)%>%
    mutate()%>%
    group_by(raceagegroupmonth)%>%
    summarise_at(vars(case, death, hosp, missingdeath, missinghosp), funs(sum), na.rm = TRUE)%>%
    ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupmonth,"_"), `[`, 1),
        agegroup=sapply(strsplit(raceagegroupmonth,"_"), `[`, 2),
        month=as.numeric(sapply(strsplit(raceagegroupmonth,"_"), `[`, 3)), 
        
        age=as.numeric(sapply(strsplit(agegroup,"-"), `[`, 1)), #age interval begins
            age=ifelse(agegroup=="80+ Years", 80, age), 
        
        raceagegroup=paste0(racegroup, "_", agegroup) 
        )%>%
    arrange(racegroup, agegroup, month)
    

    #View(dtamonthly)
    #nrow(dtamonthly)
    #str(dtamonthly)
    #summary(dtamonthly)

dim(dtamonthly)
dtamonthly<-left_join(dtamonthly, dtapopagerace10, by = c("racegroup", "age")) 
dim(dtamonthly)    

write.csv(dtamonthly, "X_COVID19_US_AgeComposition/dtamonthly_20201204.csv")
```

```{r dtaweekly}
# Weekly trend count data by RACE AND AGE
dtaweekly<-dtacase%>%
    arrange(racegroup, agegroup, week)%>%
    mutate(raceagegroupweek=paste0(racegroup,"_",agegroup,"_",as.character(week)))%>%
    group_by(raceagegroupweek)%>%
    summarise_at(vars(case, death, hosp, missingdeath, missinghosp), funs(sum), na.rm = TRUE)%>%
    ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 1),
        agegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 2),
        week=as.numeric(sapply(strsplit(raceagegroupweek,"_"), `[`, 3)), 
        
        age=as.numeric(sapply(strsplit(agegroup,"-"), `[`, 1)), #age interval begins
            age=ifelse(agegroup=="80+ Years", 80, age), 
        
        raceagegroup=paste0(racegroup, "_", agegroup) 
        )%>%
    arrange(racegroup, agegroup, week)%>%
    mutate(
        case_lag1 =lag(case), 
        case_lag2 =lag(case_lag1), 
        case_lag3 =lag(case_lag2), 
            test1=raceagegroup!=lag(raceagegroup),
            test2=lag(test1)==TRUE, 
            test3=lag(test2)==TRUE, 
            case_lag1 = ifelse(test1==TRUE, NA, case_lag1),
            case_lag2 = ifelse(test1==TRUE, NA, case_lag2),    
            case_lag2 = ifelse(test2==TRUE, NA, case_lag2),
            case_lag3 = ifelse(test1==TRUE, NA, case_lag3),    
            case_lag3 = ifelse(test2==TRUE, NA, case_lag3),
            case_lag3 = ifelse(test3==TRUE, NA, case_lag3),
        #cfr_lag=round(100*death/(case+case_lag1+case_lag2), 2), 
        cfr=round(100*death/case, 2) 
        )%>%
    select(-test1, -test2, -test3)

    #View(dtaweekly)
    #nrow(dtaweekly)
    #str(dtaweekly)
    #summary(dtaweekly)

dim(dtaweekly)
dtaweekly<-left_join(dtaweekly, dtapopagerace10, by = c("racegroup", "age")) 
dim(dtaweekly)    

write.csv(dtaweekly, "X_COVID19_US_AgeComposition/dtaweekly_20201204.csv")
```

```{r dtaweeklycum}
# Weekly trend count data by RACE AND AGE
dtaweeklycum<-dtaweekly%>%
    arrange(racegroup, agegroup, week)%>%
    group_by(raceagegroup)%>%
    mutate(
        cumcase =cumsum(case),   
        cumdeath=cumsum(death), 
        cumhosp =cumsum(hosp)
    )%>%
    ungroup()%>%
    arrange(racegroup, agegroup, week)

colnames(dtaweeklycum)

    #nrow(dtaweekly)
    #str(dtaweekly)
    #summary(dtaweekly)

```

##__Age distribution of population, confirmed COVID-19 cases, and COVID-19 deaths in the US__ 

```{r}
totalcases<-as.character(sum(dtacase$case))
totaldeaths<-as.character(table(dtacase$death)[[2]])
latestweek<-max(dtacase$week)
latestdate<-as.character(as.Date.POSIXct(max(dtacase$cdc_report_dt)))

missingage  <-round(100*(prop.table(table(dtacase$missingage))[[2]]), 1)
missingsex <-round(100*(prop.table(table(dtacase$missingsex))[[2]]), 1)
missingrace <-round(100*(prop.table(table(dtacase$missingrace))[[2]]), 1)
missingdeath<-round(100*(prop.table(table(dtacase$missingdeath))[[2]]), 1)
missinghosp <-round(100*(prop.table(table(dtacase$missinghosp))[[2]]), 1)

temp<-dtacase%>%filter(death==1)
missingagedeath <-round(100*(prop.table(table(temp$missingage))[[2]]), 3)
missingsexdeath <-round(100*(prop.table(table(temp$missingsex))[[2]]), 1)
missingracedeath <-round(100*(prop.table(table(temp$missingrace))[[2]]), 1)

```

(Updated: `r time` EDT)  

* Data source: [COVID-19 Case Surveillance Public Use Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf)     
* Accessed on December 25, 2020 (version released on December 4, 2020)   
* Latest report week is `r latestweek`, and latest report date is November 19, 2020.    
* Total number of cases included in the data source: `r totalcases`    
* Total number of deaths included in the data source: `r totaldeaths`    
* NOTE: 
    1. Age reporting is good. Missing in only `r missingage`% of cases and `r missingagedeath`% of deaths.   
    2. Sex reporting is also good. Missing in `r missingsex`% of cases and `r missingsexdeath`% of deaths.   
    3. Race reporting is missing in many cases (`r missingrace`%). Reporting among reported deaths is better: `r missingracedeath`% of reported deaths have missing race.  
    4. Survival status and hospitalizations is missing in `r missingdeath`% and `r missinghosp`% of cases, respectively.    

```{r}
#* Latest report week is `r latestweek`, and latest report date is `r latestdate`. 
#* Total number of cases included in the data source: `r totalcases`    
#* Total number of deaths included in the data source: `r totaldeaths`    
#* NOTE: 
#    1. Age reporting is good. Missing in only `r missingage`% of cases and `r missingagedeath`% of deaths.   
#    2. Sex reporting is also good. Missing in `r missingsex`% of cases and `r missingsexdeath`% of deaths.   
#    3. Race reporting is missing in many cases (`r missingrace`%). Reporting among reported deaths is better: `r missingracedeath`% of reported deaths have missing race.  
#    4. Survival status and hospitalizations is missing in `r missingdeath`% and `r missinghosp`% of cases, respectively.    
```
##__1. Overall age distribution__ 

```{r overall_plotagecompositionDTA}
#Population

temppop <- dtaweekly%>%
    filter(week==max(week))%>%    
    select(agegroup, pop)%>%
    filter(is.na(pop)==FALSE)%>%
    group_by(agegroup)%>%summarize_all(funs(sum))%>%
    mutate(
        total=sum(pop, na.rm = TRUE), 
        pct=round(100*pop/total,1)  ,
        group="1. Population"  
    )%>%
    select(agegroup, group, total, pct)

#cases
tempcase <- dtaweekly%>%select(agegroup, case)%>%
    group_by(agegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=sum(case), 
        pct=round(100*case/total,1),
        group="2. Cases"
    )%>%
    select(-case)

#deaths
tempdeath <- dtaweekly%>%select(agegroup, death)%>%
    group_by(agegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=sum(death), 
        pct=round(100*death/total,1),
        group="3. Deaths (Reported)"
    )%>%
    select(-death)

# merge all three
dtafig<-rbind(temppop, tempcase, tempdeath) 
dim(dtafig)
str(dtafig)
```

####__1.1. Population, cumulative cases, and cumulative deaths__
```{r overall_plotagecomposition, results="asis", results="asis", fig.align="left", out.width="800px", out.height="300px"}
dtafig%>%
    plot_ly(x=~group,
            y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    layout(
        title ="",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

####__1.2. Case trend__
```{r overall_plotagecomposition_trendDTA}

temp1 <- dtaweekly%>%select(week, case)%>%
    group_by(week)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=case
    )%>%
    select(-case)

temp2 <- dtaweekly%>%
    mutate(agegroupweek=paste0(agegroup,"_",as.character(week)))%>%
    select(agegroupweek, case)%>%
    group_by(agegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        agegroup=sapply(strsplit(agegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(agegroupweek,"_"), `[`, 2))
    )

    dim(temp1)
    dim(temp2)

dtafig<-left_join(temp2, temp1, by = "week")%>% 
    mutate(
        pct=round(100*case/total,1)
    )

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day

sum(dtacase$case)
sum(dtaweekly$case)
sum(dtafig$case)
```

#####__1.2.1. Weekly reported cases, number__
```{r overall_plotagecomposition_trend1, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig%>%
    plot_ly(x=~weekdate,
            y = ~case, type = "bar",
            color= ~agegroup ,
            colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Number of reported cases per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

#####__1.2.2. Weekly reported cases, percent__
(only weeks with more than 20 cases)
```{r overall_plotagecomposition_trend2, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig%>%
    filter(total>20)%>%
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
            color= ~agegroup ,
            colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Percent of reported cases per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

####__1.3. hospitalizations trend__
```{r overall_plotagecomposition_hosp_trendDTA}

temp1 <- dtaweekly%>%select(week, hosp)%>%
    group_by(week)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=hosp
    )%>%
    select(-hosp)

temp2 <- dtaweekly%>%
    mutate(agegroupweek=paste0(agegroup,"_",as.character(week)))%>%
    select(agegroupweek, hosp)%>%
    group_by(agegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        agegroup=sapply(strsplit(agegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(agegroupweek,"_"), `[`, 2))
    )

    dim(temp1)
    dim(temp2)

dtafig<-left_join(temp2, temp1, by = "week")%>% 
    mutate(
        pct=round(100*hosp/total,1)
    )

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day


sum(dtacase$hosp)
sum(dtaweekly$hosp)
sum(dtafig$hosp)
```

#####__1.3.1. Weekly reported hospitalizations, number__
```{r overall_plotagecomposition_hosp_trend1, results="asis", fig.align="left", out.width="800px", out.height="300px"}


dtafig%>%
    plot_ly(x=~weekdate,
            y = ~hosp, type = "bar",
            color= ~agegroup ,
            colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Number of reported hospitalizations per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__1.3.2. Weekly reported hospitalizations, percent__
(only weeks with more than 20 hospitalizations)
```{r overall_plotagecomposition_hosp_trend2, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig%>%
    filter(total>20)%>%
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
            color= ~agegroup ,
            colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Percent of reported hospitalizations per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

####__1.4. Deaths trend__
```{r overall_plotagecomposition_death_trendDTA}

temp1 <- dtaweekly%>%select(week, death)%>%
    group_by(week)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=death
    )%>%
    select(-death)

temp2 <- dtaweekly%>%
    mutate(agegroupweek=paste0(agegroup,"_",as.character(week)))%>%
    select(agegroupweek, death)%>%
    group_by(agegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        agegroup=sapply(strsplit(agegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(agegroupweek,"_"), `[`, 2))
    )

    dim(temp1)
    dim(temp2)

dtafig<-left_join(temp2, temp1, by = "week")%>% 
    mutate(
        pct=round(100*death/total,1)
    )

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day

sum(dtacase$death)
sum(dtaweekly$death)
sum(dtafig$death)
```

#####__1.4.1. Weekly reported deaths, number__
```{r overall_plotagecomposition_death_trend1, results="asis", fig.align="left", out.width="800px", out.height="300px"}


dtafig%>%
    plot_ly(x=~weekdate,
            y = ~death, type = "bar",
            color= ~agegroup ,
            colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Number of reported deaths per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__1.4.2. Weekly reported deaths, percent__
(only weeks with more than 20 deaths)
```{r overall_plotagecomposition_death_trend2, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig%>%
    filter(total>20)%>%
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
            color= ~agegroup ,
            colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Percent of reported deaths per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

##__2. Overall race distribution__ 

```{r overall_plotracecompositionDTA}
#Population

temppop <- dtaweekly%>%
    filter(week==max(week))%>%    
    select(racegroup, pop)%>%
    filter(is.na(pop)==FALSE)%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%
    mutate(
        total=sum(pop, na.rm = TRUE), 
        pct=round(100*pop/total,1)  ,
        group="1. Population"  
    )%>%
    select(racegroup, group, total, pct)

#cases
tempcase <- dtaweekly%>%select(racegroup, case)%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=sum(case), 
        pct=round(100*case/total,1),
        group="2. Cases"
    )%>%
    select(-case)

#deaths
tempdeath <- dtaweekly%>%select(racegroup, death)%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=sum(death), 
        pct=round(100*death/total,1),
        group="3. Deaths (Reported)"
    )%>%
    select(-death)

# merge all three
dtafig<-rbind(temppop, tempcase, tempdeath) 
dim(dtafig)
str(dtafig)
```

####__2.1. Population, cumulative cases, and cumulative deaths__
```{r overall_plotracecomposition, results="asis", fig.align="left", out.width="800px", out.height="300px"}
# color ccheme: single hue 9-scale gree & blue
### https://colorbrewer2.org/?type=sequential&scheme=YlOrRd&n=9

fig<-dtafig%>%
    plot_ly(x=~group,
            y = ~pct, type = "bar", 
            color=~racegroup
            ) %>% 
    layout(
        title ="",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

dtafig%>%
    plot_ly(x=~group,
            y = ~pct, type = "bar", 
            color = ~racegroup, 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6',
                       'lightgray')
            ) %>% 
    layout(
        title ="",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

####__2.2. Case trend__
```{r overall_plotracecomposition_trendDTA}

temp1 <- dtaweekly%>%select(week, case)%>%
    group_by(week)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=case
    )%>%
    select(-case)

temp2 <- dtaweekly%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, case)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

    dim(temp1)
    dim(temp2)

dtafig<-left_join(temp2, temp1, by = "week")%>% 
    mutate(
        pct=round(100*case/total,1)
    )

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day


sum(dtacase$case)
sum(dtaweekly$case)
sum(dtafig$case)
```

#####__2.2.1. Weekly reported cases, number__
```{r overall_plotracecomposition_trend1, results="asis", fig.align="left", out.width="800px", out.height="300px"}


dtafig%>%
    plot_ly(x=~weekdate,
            y = ~case, type = "bar",
            color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6',
                       'lightgray')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Number of reported cases per week",
                     titlefont=list(size=12)), 
        xaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

#####__2.2.2. Weekly reported cases, percent__
(only weeks with more than 20 cases)
```{r overall_plotracecomposition_trend2, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig1<-dtafig%>%
    filter(total>20)%>%
    select(weekdate, pct, racegroup)

fig1<-dtafig1%>%    
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
            color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6',
                       'lightgray')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Percent of reported cases per week",
                     titlefont=list(size=12)), 
        xaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

fig1

```

####__2.3. Deaths trend__
```{r overall_plotracecomposition_death_trendDTA}

temp1 <- dtaweekly%>%select(week, death)%>%
    group_by(week)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=death
    )%>%
    select(-death)

temp2 <- dtaweekly%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, death)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

    dim(temp1)
    dim(temp2)

dtafig<-left_join(temp2, temp1, by = "week")%>% 
    mutate(
        pct=round(100*death/total,1)
    )

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day


sum(dtacase$death)
sum(dtaweekly$death)
sum(dtafig$death)
```

#####__2.3.1. Weekly reported deaths, number__
```{r overall_plotracecomposition_death_trend1, results="asis", fig.align="left", out.width="800px", out.height="300px"}


dtafig%>%
    plot_ly(x=~weekdate,
            y = ~death, type = "bar",
            color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6',
                       'lightgray')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Number of reported deaths per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__2.3.2. Weekly reported deaths, percent__
(only weeks with more than 20 deaths)
```{r overall_plotracecomposition_death_trend2, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig%>%
    filter(total>20)%>%
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
            color= ~racegroup, 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6',
                       'lightgray')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Percent of reported deaths per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

##__2.A. Only among observations with known race: Overall race distribution__ 
```{r overall_plotracecompositionDTA_knownrace}
#Population

temppop <- dtaweekly%>%
    filter(week==max(week))%>%    
    select(racegroup, pop)%>%
    filter(racegroup!="Unknown")%>%
    filter(is.na(pop)==FALSE)%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%
    mutate(
        total=sum(pop, na.rm = TRUE), 
        pct=round(100*pop/total,1)  ,
        group="1. Population"  
    )%>%
    select(racegroup, group, total, pct)

#cases
tempcase <- dtaweekly%>%select(racegroup, case)%>%
    filter(racegroup!="Unknown")%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=sum(case), 
        pct=round(100*case/total,1),
        group="2. Cases"
    )%>%
    select(-case)

#deaths
tempdeath <- dtaweekly%>%select(racegroup, death)%>%
    filter(racegroup!="Unknown")%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=sum(death), 
        pct=round(100*death/total,1),
        group="3. Deaths (Reported)"
    )%>%
    select(-death)

# merge all three
dtafig<-rbind(temppop, tempcase, tempdeath) 
dim(dtafig)
str(dtafig)
```

####__2.A.1. Population, cumulative cases, and cumulative deaths__
```{r overall_plotracecomposition_knownrace, results="asis", fig.align="left", out.width="800px", out.height="300px"}
dtafig%>%
    plot_ly(x=~group,
            y = ~pct, type = "bar",
             color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6')
             ) %>% 
    layout(
        title ="",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

####__2.A.2. Case trend__
```{r overall_plotracecomposition_case_trendDTA_knownrace}

temp1 <- dtaweekly%>%
    filter(racegroup!="Unknown")%>%
    select(week, case)%>%
    group_by(week)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=case
    )%>%
    select(-case)

temp2 <- dtaweekly%>%
    filter(racegroup!="Unknown")%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, case)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

    dim(temp1)
    dim(temp2)

dtafig<-left_join(temp2, temp1, by = "week")%>% 
    mutate(
        pct=round(100*case/total,1)
    )

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day


sum(dtacase$case)
sum(dtaweekly$case)
sum(dtafig$case)
```

#####__2.A.2.1. Weekly reported cases, number__
```{r overall_plotracecomposition_case_trend1_knownrace, results="asis", fig.align="left", out.width="800px", out.height="300px"}


dtafig%>%
    plot_ly(x=~weekdate,
            y = ~case, type = "bar",
            color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Number of reported cases per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__2.A.2.2. Weekly reported cases, percent__
(only weeks with more than 20 cases)
```{r overall_plotracecomposition_case_trend2_knownrace, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig%>%
    filter(total>20)%>%
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
            color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Percent of reported cases per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

####__2.A.3. hosp trend__
```{r overall_plotracecomposition_hosp_trendDTA_knownrace}

temp1 <- dtaweekly%>%
    filter(racegroup!="Unknown")%>%
    select(week, hosp)%>%
    group_by(week)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=hosp
    )%>%
    select(-hosp)

temp2 <- dtaweekly%>%
    filter(racegroup!="Unknown")%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, hosp)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

    dim(temp1)
    dim(temp2)

dtafig<-left_join(temp2, temp1, by = "week")%>% 
    mutate(
        pct=round(100*hosp/total,1)
    )

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day


sum(dtacase$hosp)
sum(dtaweekly$hosp)
sum(dtafig$hosp)
```

#####__2.A.3.1. Weekly reported hospitlizations, number__
```{r overall_plotracecomposition_hosp_trend1_knownrace, results="asis", fig.align="left", out.width="800px", out.height="300px"}


dtafig%>%
    plot_ly(x=~weekdate,
            y = ~hosp, type = "bar",
            color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Number of reported hospitlizations per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__2.A.3.2. Weekly reported hospitlizations, percent__
(only weeks with more than 20 hospitalizations)
```{r overall_plotracecomposition_hosp_trend2_knownrace, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig%>%
    filter(total>20)%>%
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
            color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Percent of reported hospitlizations per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

####__2.A.4. Deaths trend__
```{r overall_plotracecomposition_death_trendDTA_knownrace}

temp1 <- dtaweekly%>%
    filter(racegroup!="Unknown")%>%
    select(week, death)%>%
    group_by(week)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=death
    )%>%
    select(-death)

temp2 <- dtaweekly%>%
    filter(racegroup!="Unknown")%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, death)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

    dim(temp1)
    dim(temp2)

dtafig<-left_join(temp2, temp1, by = "week")%>% 
    mutate(
        pct=round(100*death/total,1)
    )

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day


sum(dtacase$death)
sum(dtaweekly$death)
sum(dtafig$death)
```

#####__2.A.4.1. Weekly reported deaths, number__
```{r overall_plotracecomposition_death_trend1_knownrace, results="asis", fig.align="left", out.width="800px", out.height="300px"}


dtafig%>%
    plot_ly(x=~weekdate,
            y = ~death, type = "bar",
            color= ~racegroup , 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Number of reported deaths per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__2.A.4.2. Weekly reported deaths, percent__
(only weeks with more than 20 deaths)
```{r overall_plotracecomposition_death_trend2_knownrace, results="asis", fig.align="left", out.width="800px", out.height="300px"}

dtafig%>%
    filter(total>20)%>%
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
            color= ~racegroup, 
            colors = c('#74c476',
                       '#e31a1c',
                       '#fed976',
                       '#feb24c',
                       '#fd8d3c',
                       '#fc4e2a',
                       '#6baed6')
            )%>% 
    layout(
        title ="",
        yaxis = list(title = "Percent of reported deaths per week",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

```

##__3. Latest cumulative age distribution by race__ 

```{r plotagecompositionDTA}
#Population

temppop <- dtaweekly%>%
    select(racegroup, agegroup, week, pop)%>%
    filter(week==max(week)) %>%    
    group_by(racegroup)%>%
    mutate(
        total=sum(pop, na.rm = TRUE), 
        pct=round(100*pop/total,1)  ,
        group="1. Population"  
    )%>%
    ungroup()%>%
    select(racegroup, agegroup, group, total, pct)%>%
    filter(agegroup!="Unknown")%>%
    filter(racegroup!="Unknown")

#cases
temp1 <- dtaweekly%>%select(racegroup, case)%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=case
    )%>%
    select(-case)

temp2 <- dtaweekly%>%select(raceagegroup, case)%>%
    group_by(raceagegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroup,"_"), `[`, 1), 
        agegroup=sapply(strsplit(raceagegroup,"_"), `[`, 2)
    )

    dim(temp1)
    dim(temp2)

tempcase<-left_join(temp2, temp1, by = "racegroup")%>% 
    mutate(
        pct=round(100*case/total,1),
        group="2. Cases"
    )%>%
    select(-case, -raceagegroup)

#hospitalizations
temp1 <- dtaweekly%>%select(racegroup, hosp)%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=hosp
    )%>%
    select(-hosp)

temp2 <- dtaweekly%>%select(raceagegroup, hosp)%>%
    group_by(raceagegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroup,"_"), `[`, 1),
        agegroup=sapply(strsplit(raceagegroup,"_"), `[`, 2)
    )

    dim(temp1)
    dim(temp2)

temphosp<-left_join(temp2, temp1, by = "racegroup")%>% 
    mutate(
        pct=round(100*hosp/total,1),
        group="3. Hospitalizations (Reported)"
    )%>%
    select(-hosp, -raceagegroup)

#deaths
temp1 <- dtaweekly%>%select(racegroup, death)%>%
    group_by(racegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=death
    )%>%
    select(-death)

temp2 <- dtaweekly%>%select(raceagegroup, death)%>%
    group_by(raceagegroup)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroup,"_"), `[`, 1),
        agegroup=sapply(strsplit(raceagegroup,"_"), `[`, 2)
    )

    dim(temp1)
    dim(temp2)

tempdeath<-left_join(temp2, temp1, by = "racegroup")%>% 
    mutate(
        pct=round(100*death/total,1),
        group="4. Deaths (Reported)"
    )%>%
    select(-death, -raceagegroup)

# merge all three
dtafig<-rbind(temppop, tempcase, temphosp, tempdeath) 
dim(dtafig)
str(dtafig)
```

####__3.1. By metric: population, cases, hopitalization, and deaths__

#####__3.1.1. Age distribution of population by race__ 
```{r plotagecomposition1-1, results="asis", results="asis", fig.align="left", out.width="800px", out.height="300px"}
dtafig%>% 
    filter(group=="1. Population")%>%
    plot_ly(x=~racegroup,
            y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    layout(
        title ="Population",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__3.1.2. Age distribution of cases by race__
```{r plotagecomposition1-2, , results="asis", fig.align="left", out.width="800px", out.height="300px"}
dtafig%>% 
    filter(group=="2. Cases")%>%
    plot_ly(x=~racegroup,
            y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    layout(
        title ="Cases",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__3.1.3. Age distribution of hospitalizations by race__
```{r plotagecomposition1-3, , results="asis", fig.align="left", out.width="800px", out.height="300px"}
dtafig%>% 
    filter(group=="3. Hospitalizations (Reported)")%>%
    plot_ly(x=~racegroup,
            y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    layout(
        title ="hospitalizations (Reported)",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__3.1.4. Age distribution of deaths by race__
```{r plotagecomposition1-4, , results="asis", fig.align="left", out.width="800px", out.height="300px"}
dtafig%>% 
    filter(group=="4. Deaths (Reported)")%>%
    plot_ly(x=~racegroup,
            y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    layout(
        title ="Deaths (Reported)",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

####__3.2. By race__
__Same data but different organizatoin__
```{r plotagecomposition2, results="asis", fig.align="left", out.width="800px", out.height="800px"}
panel <- . %>% 
    plot_ly(x=~group,
            y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    add_annotations(
        text = ~unique(racegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%    
    layout(
        title ="",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

dtafig$group<-fct_relevel(factor(dtafig$group, levels = unique(dtafig$group)),
                                  c("1. Population", 
                                    "2. Cases", 
                                    "3. Hospitalizations (Reported)", 
                                    "4. Deaths (Reported)"))

dtafig%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Percent"), 
        showlegend = FALSE
        )  

```

##__4. Trends of age distribution by race__ 

```{r plotagecomposition_trendDTA}

#cases
temp1 <- dtaweeklycum%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, cumcase)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=cumcase
    )%>%
    select(-cumcase)%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

temp2 <- dtaweeklycum%>%
    select(raceagegroupweek, cumcase)%>%
    group_by(raceagegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 1), 
        agegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 2),
        week=as.numeric(sapply(strsplit(raceagegroupweek,"_"), `[`, 3))
    )

    dim(temp1)
    dim(temp2)
    colnames(temp1)
    colnames(temp2)

tempcase<-left_join(temp2, temp1, by = c("racegroup", "week"))%>% 
    mutate(
        pct=round(100*cumcase/total,1),
        group="Cases",
        count=cumcase
    )%>%
    select(raceagegroupweek, racegroup, agegroup, week, total, pct, count, group) 

#hospitalizations
#hosp
temp1 <- dtaweeklycum%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, cumhosp)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=cumhosp
    )%>%
    select(-cumhosp)%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

temp2 <- dtaweeklycum%>%
    select(raceagegroupweek, cumhosp)%>%
    group_by(raceagegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 1), 
        agegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 2),
        week=as.numeric(sapply(strsplit(raceagegroupweek,"_"), `[`, 3))
    )

    dim(temp1)
    dim(temp2)
    colnames(temp1)
    colnames(temp2)

temphosp<-left_join(temp2, temp1, by = c("racegroup", "week"))%>% 
    mutate(
        pct=round(100*cumhosp/total,1),
        group="hospitalizations",
        count=cumhosp
    )%>%
    select(raceagegroupweek, racegroup, agegroup, week, total, pct, count, group) 

#death
temp1 <- dtaweeklycum%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, cumdeath)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=cumdeath
    )%>%
    select(-cumdeath)%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

temp2 <- dtaweeklycum%>%
    select(raceagegroupweek, cumdeath)%>%
    group_by(raceagegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 1), 
        agegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 2),
        week=as.numeric(sapply(strsplit(raceagegroupweek,"_"), `[`, 3))
    )

    dim(temp1)
    dim(temp2)
    colnames(temp1)
    colnames(temp2)

tempdeath<-left_join(temp2, temp1, by = c("racegroup", "week"))%>% 
    mutate(
        pct=round(100*cumdeath/total,1), 
        group="Deaths", 
        count=cumdeath
    )%>%
    select(raceagegroupweek, racegroup, agegroup, week, total, pct, count, group) 

# merge all three
dtafig<-rbind(tempcase, temphosp, tempdeath) 

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day

dim(dtafig)
str(dtafig)
```

####__4.1 Cumulative percent __

```{r plotagecomposition_trend_pct}
panel <- . %>% 
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    add_annotations(
        text = ~unique(racegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%    
    layout(
        title ="",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__4.1.1. Trends of cumulative percent in cases__
(only weeks with more than 20 cases)
```{r plotagecomposition_trend_case_pct, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Cases")%>%
    filter(total>20)%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Percent"), 
        showlegend = FALSE
        )  

```

#####__4.1.2. Trends of cumulative percent in hospitalizations__
(only weeks with more than 20 hospitalizations)
```{r plotagecomposition_trend_hosp_pct, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="hospitalizations")%>%
    filter(total>20)%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Percent"), 
        showlegend = FALSE
        )  
```

#####__4.1.3. Trends of cumulative percent in deaths__
(only weeks with more than 20 deaths)
```{r plotagecomposition_trend_death_pct, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Deaths")%>%
    filter(total>20)%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Percent"), 
        showlegend = FALSE
        )  
```

####__4.2 Cumulative count __

```{r plotagecomposition_trend_num}
panel <- . %>% 
    plot_ly(x=~weekdate,
            y = ~count, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    add_annotations(
        text = ~unique(racegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%    
    layout(
        title ="",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__4.2.1. Trends of cumulative count in cases__
```{r plotagecomposition_trend_case_num, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Cases")%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Percent"), 
        showlegend = FALSE
        )  
```

#####__4.2.2. Trends of cumulative count in hospitalizations__
```{r plotagecomposition_trend_hosp_num, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="hospitalizations")%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Percent"), 
        showlegend = FALSE
        )  
```

#####__4.2.3. Trends of cumulative count in deaths__
```{r plotagecomposition_trend_death_num, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Deaths")%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Percent"), 
        showlegend = FALSE
        )  
```


####__4.3 Weekly percent __

```{r plotagecomposition_weeklytrendDTA}

#cases
temp1 <- dtaweekly%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, case)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=case
    )%>%
    select(-case)%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

temp2 <- dtaweekly%>%
    select(raceagegroupweek, case)%>%
    group_by(raceagegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 1), 
        agegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 2),
        week=as.numeric(sapply(strsplit(raceagegroupweek,"_"), `[`, 3))
    )

    dim(temp1)
    dim(temp2)
    colnames(temp1)
    colnames(temp2)

tempcase<-left_join(temp2, temp1, by = c("racegroup", "week"))%>% 
    mutate(
        pct=round(100*case/total,1),
        group="Cases",
        count=case
    )%>%
    select(raceagegroupweek, racegroup, agegroup, week, total, pct, count, group) 

#hospitalizations
#hosp
temp1 <- dtaweekly%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, hosp)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=hosp
    )%>%
    select(-hosp)%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

temp2 <- dtaweekly%>%
    select(raceagegroupweek, hosp)%>%
    group_by(raceagegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 1), 
        agegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 2),
        week=as.numeric(sapply(strsplit(raceagegroupweek,"_"), `[`, 3))
    )

    dim(temp1)
    dim(temp2)
    colnames(temp1)
    colnames(temp2)

temphosp<-left_join(temp2, temp1, by = c("racegroup", "week"))%>% 
    mutate(
        pct=round(100*hosp/total,1),
        group="hospitalizations",
        count=hosp
    )%>%
    select(raceagegroupweek, racegroup, agegroup, week, total, pct, count, group) 

#death
temp1 <- dtaweekly%>%
    mutate(racegroupweek=paste0(racegroup,"_",as.character(week)))%>%
    select(racegroupweek, death)%>%
    group_by(racegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        total=death
    )%>%
    select(-death)%>%
    mutate(
        racegroup=sapply(strsplit(racegroupweek,"_"), `[`, 1), 
        week=as.numeric(sapply(strsplit(racegroupweek,"_"), `[`, 2))
    )

temp2 <- dtaweekly%>%
    select(raceagegroupweek, death)%>%
    group_by(raceagegroupweek)%>%summarize_all(funs(sum))%>%ungroup()%>%
    mutate(
        racegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 1), 
        agegroup=sapply(strsplit(raceagegroupweek,"_"), `[`, 2),
        week=as.numeric(sapply(strsplit(raceagegroupweek,"_"), `[`, 3))
    )

    dim(temp1)
    dim(temp2)
    colnames(temp1)
    colnames(temp2)

tempdeath<-left_join(temp2, temp1, by = c("racegroup", "week"))%>% 
    mutate(
        pct=round(100*death/total,1), 
        group="Deaths", 
        count=death
    )%>%
    select(raceagegroupweek, racegroup, agegroup, week, total, pct, count, group) 

# merge all three
dtafig<-rbind(tempcase, temphosp, tempdeath) 

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))
dtafig$weekdate<-as.Date("2020-01-01") + dtafig$week * 7 - first.day

dim(dtafig)
str(dtafig)
```

```{r plotagecomposition_weeklytrend_pct}
panel <- . %>% 
    plot_ly(x=~weekdate,
            y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    add_annotations(
        text = ~unique(racegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%    
    layout(
        title ="",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        xaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__4.3.1. Trends of weekly percent in cases__
(only weeks with more than 20 cases)
```{r plotagecomposition_weeklytrend_case_pct, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Cases")%>%
    filter(total>20)%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "% of weekly cases"), 
        xaxis=list(title = ""), 
        showlegend = FALSE
        )  

dtafig2case<-dtafig%>%
    filter(group=="Cases")%>%
    filter(total>20)%>%    
    filter(racegroup=="Hispanic" | racegroup=="NH Black" | racegroup=="NH White")%>%
    select(weekdate, pct, agegroup, racegroup)

fig2case<-dtafig2case%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 1, shareX = TRUE, shareY = TRUE)%>%
    layout(
        yaxis=list(title = "% of weekly cases"), 
        xaxis=list(title = ""), 
        showlegend = FALSE
        ) 


```

#####__4.3.2. Trends of weekly percent in hospitalizations__
(only weeks with more than 20 hospitalizations)
```{r plotagecomposition_weeklytrend_hosp_pct, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="hospitalizations")%>%
    filter(total>20)%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "% of weekly hospitalizations"), 
        xaxis=list(title = ""), 
        showlegend = FALSE
        )  


dtafig2hosp<-dtafig%>%
    filter(group=="hospitalizations")%>%
    filter(total>20)%>%
    filter(racegroup=="Hispanic" | racegroup=="NH Black" | racegroup=="NH White")%>%
    select(weekdate, pct, agegroup, racegroup)

fig2hosp<-dtafig2hosp%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 1, shareX = TRUE, shareY = TRUE)%>%
    layout(
        yaxis=list(title = "% of weekly hospitalizations"), 
        xaxis=list(title = ""), 
        showlegend = FALSE
        )  
```

#####__4.3.3. Trends of weekly percent in deaths__
(only weeks with more than 20 deaths)
```{r plotagecomposition_weeklytrend_death_pct, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Deaths")%>%
    filter(total>20)%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "% of weekly deaths"), 
        xaxis=list(title = ""), 
        showlegend = FALSE
        )  

dtafig2death<-dtafig%>%
    filter(group=="Deaths")%>%
    filter(total>20)%>%    
    filter(racegroup=="Hispanic" | racegroup=="NH Black" | racegroup=="NH White")%>%
    select(weekdate, pct, agegroup, racegroup)

fig2death<-dtafig2death%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 1, shareX = TRUE, shareY = TRUE)%>%
    layout(
        yaxis=list(title = "% of weekly deaths"), 
        xaxis=list(title = ""), 
        showlegend = FALSE
        )  
```

(only weeks with more than 30 deaths)
```{r plotagecomposition_weeklytrend_death_pct2, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Deaths")%>%
    filter(total>30)%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Percent"), 
        showlegend = FALSE
        )  
```

####__4.4 weekly number __

```{r plotagecomposition_weeklytrend_count}
panel <- . %>% 
    plot_ly(x=~weekdate,
            y = ~count, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    add_annotations(
        text = ~unique(racegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%    
    layout(
        title ="",
        yaxis = list(title = "Number",
                     titlefont=list(size=12)), 
        xaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 
```

#####__4.4.1. Trends of weekly number in cases__
```{r plotagecomposition_weeklytrend_case_count, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Cases")%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Number of cases"), 
        showlegend = FALSE
        )  
```

#####__4.4.2. Trends of weekly number in hospitalizations__
```{r plotagecomposition_weeklytrend_hosp_count, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="hospitalizations")%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Number of hospitalizations"), 
        showlegend = FALSE
        )  
```

#####__4.4.3. Trends of weekly number in deaths__
```{r plotagecomposition_weeklytrend_death_count, results="asis", fig.align="left", out.width="800px", out.height="800px"}
dtafig%>%
    filter(group=="Deaths")%>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 4, shareX = TRUE, shareY = FALSE, margin=0.05)%>%
    layout(
        yaxis=list(title = "Number of deaths"), 
        showlegend = FALSE
        )  
```

####__4.5. Trends of weekly lagged CFR__

```{r cfrbyage}

panel <- . %>% 
    plot_ly(x=~weekdate, y=~cfr80, name = "80+", 
               type = 'scatter', mode='lines',
               line = list(color = c( '#3288bd')) )%>% 
    add_trace(y=~cfr70, name = "70-79", line = list(color = c( '#66c2a5'))) %>% 
    add_trace(y=~cfr60, name = "60-69", line = list(color = c( '#abdda4'))) %>% 
    add_trace(y=~cfr50, name = "50-59", line = list(color = c( '#e6f598'))) %>% 
    add_trace(y=~cfr40, name = "40-49", line = list(color = c( '#ffffbf'))) %>% 
    add_trace(y=~cfr30, name = "30-39", line = list(color = c( '#fee08b'))) %>% 
    add_trace(y=~cfr20, name = "20-29", line = list(color = c( '#fdae61'))) %>% 
    add_trace(y=~cfr10, name = "10-19", line = list(color = c( '#f46d43'))) %>% 
    add_trace(y=~cfr0,  name = "0-9",   line = list(color = c( '#d53e4f'))) %>% 
    add_annotations(
        text = ~unique(racegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%     
    layout(
        yaxis = list(type = "log", range = c(-2, 2) ), 
        xaxis = list(title = "", titlefont=list(size=8), tickfont=list(size=8)),
        legend = list(font=list(size=10)) 
        ) 

first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))

dtasortedbyage<-dtaweekly %>%
    filter(is.na(age)==FALSE)%>% #drop cases with missing cases
    filter(week>=14)%>% # drop initial weeks
    filter(racegroup!="Unknown")%>%  # drop cases with unknown race
    select(week, age, racegroup, 
           starts_with("case"), starts_with("death"), starts_with("cfr")) %>%
    mutate(
        weekdate=as.Date("2020-01-01") + week * 7 - first.day, 
        age=paste0("cfr",age)
    ) %>%
    arrange(racegroup, weekdate, age)
```

__2-week lag: deaths in week X / cases in weeks X-1 &  X-2 (%)__ 
```{r plotcfr, results="asis", fig.align="left", out.width="800px", out.height="800px"}

dtasortedbyage %>% 
    select(racegroup, weekdate, age, cfr)%>%
    spread(age, cfr, fill = NA, convert = FALSE) %>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 3, shareX = TRUE, shareY = TRUE, margin=0.05)%>%
    layout(
        showlegend = FALSE,
        title = "Deaths / 2-week lagged cases (per 100)",
        yaxis = list(title = "deaths / 2-week lagged cases (per 100)",
                     titlefont=list(size=12),
                     type = "log", range = c(-2, 2) )
        )  
```

```{r cfrbyrace}
            
panelbyraceall <- . %>% 
    plot_ly(x=~weekdate, y=~cfrHispanic, name = "Hispanic", 
               type = 'scatter', mode='lines',
               line = list(color = c( '#74c476')) ) %>% 
    add_trace(y = ~cfrNHAIAN, name = "NH AIAN", line = list(color = c( '#e31a1c'))) %>% 
    add_trace(y = ~cfrNHAsian, name = "NH Asian", line = list(color = c( '#fed976'))) %>% 
    add_trace(y = ~cfrNHBlack, name = "NH Black", line = list(color = c( '#feb24c'))) %>% 
    add_trace(y = ~cfrNHAIAN, name = "NH Multiple/Other", line = list(color = c( '#fd8d3c'))) %>% 
    add_trace(y = ~cfrNHAsian, name = "NH NHPI", line = list(color = c( '#fc4e2a'))) %>% 
    add_trace(y = ~cfrNHWhite, name = "NH White", line = list(color = c( '#6baed6'))) %>% 
    add_annotations(
        text = ~unique(agegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%     
    layout(
        yaxis = list(type = "log", range = c(-2, 2) ), 
        xaxis = list(title = "", titlefont=list(size=8), tickfont=list(size=8)),
        legend = list(font=list(size=10)) 
        ) 

panelbyrace <- . %>% 
    plot_ly(x=~weekdate, y=~cfrHispanic, name = "Hispanic", 
               type = 'scatter', mode='lines',
               line = list(color = c( '#74c476')) ) %>% 
    add_trace(y = ~cfrNHBlack, name = "NH Black", line = list(color = c( '#feb24c'))) %>% 
    add_trace(y = ~cfrNHWhite, name = "NH White", line = list(color = c( '#6baed6'))) %>% 
    add_annotations(
        text = ~unique(agegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%     
    layout(
        yaxis = list(type = "log", range = c(-2, 2) ), 
        xaxis = list(title = "", titlefont=list(size=8), tickfont=list(size=8)),
        legend = list(font=list(size=10)) 
        ) 

panelbyraceall_linear <- . %>% 
    plot_ly(x=~weekdate, y=~cfrHispanic, name = "Hispanic", 
               type = 'scatter', mode='lines',
               line = list(color = c( '#74c476')) ) %>% 
    add_trace(y = ~cfrNHAIAN, name = "NH AIAN", line = list(color = c( '#e31a1c'))) %>% 
    add_trace(y = ~cfrNHAsian, name = "NH Asian", line = list(color = c( '#fed976'))) %>% 
    add_trace(y = ~cfrNHBlack, name = "NH Black", line = list(color = c( '#feb24c'))) %>% 
    add_trace(y = ~cfrNHAIAN, name = "NH Multiple/Other", line = list(color = c( '#fd8d3c'))) %>% 
    add_trace(y = ~cfrNHAsian, name = "NH NHPI", line = list(color = c( '#fc4e2a'))) %>% 
    add_trace(y = ~cfrNHWhite, name = "NH White", line = list(color = c( '#6baed6'))) %>% 
    add_annotations(
        text = ~unique(agegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%     
    layout(
        xaxis = list(title = "", titlefont=list(size=8), tickfont=list(size=8)),
        legend = list(font=list(size=10)) 
        ) 

panelbyrace_linear <- . %>% 
    plot_ly(x=~weekdate, y=~cfrHispanic, name = "Hispanic", 
               type = 'scatter', mode='lines',
               line = list(color = c( '#74c476')) ) %>% 
    add_trace(y = ~cfrNHBlack, name = "NH Black", line = list(color = c( '#feb24c'))) %>% 
    add_trace(y = ~cfrNHWhite, name = "NH White", line = list(color = c( '#6baed6'))) %>% 
    add_annotations(
        text = ~unique(agegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%     
    layout(
        xaxis = list(title = "", titlefont=list(size=8), tickfont=list(size=8)),
        legend = list(font=list(size=10)) 
        ) 
first.day <- as.numeric(format(as.Date("2020-01-01"), "%w"))

dtasortedbyrace<-dtaweekly %>%
    filter(is.na(age)==FALSE)%>% #drop cases with missing cases
    filter(week>=14)%>% # drop initial weeks
    filter(racegroup!="Unknown")%>%  # drop cases with unknown race
    select(week, agegroup, racegroup, 
           starts_with("case"), starts_with("death"), starts_with("cfr")) %>%
    mutate(
        weekdate=as.Date("2020-01-01") + week * 7 - first.day, 
        racegroup=paste0("cfr",racegroup)
    ) %>%
    arrange(agegroup, weekdate, racegroup)
```

```{r plotcfr_byrace, results="asis", fig.align="left", out.width="800px", out.height="800px"}

dtasortedbyrace %>% 
    select(racegroup, weekdate, agegroup, cfr)%>%
    
    mutate(racegroup=gsub("\\s", "", racegroup))%>%
    mutate(agegroup=gsub(" Years", "", agegroup))%>%
    spread(racegroup, cfr, fill = NA, convert = FALSE) %>%
    group_by(agegroup) %>%
    do(p = panelbyraceall(.)) %>%
    subplot(nrows = 3, shareX = TRUE, shareY = FALSE)%>%
    layout(
        showlegend = TRUE,
        yaxis = list(title = "deaths / 2-week lagged cases (per 100)",
                     titlefont=list(size=12),
                     type = "log", range = c(-2, 2) )
        )  
```

## Figure 1. Trend of race-distribution in weekly COVID-19 cases

```{r fig1, results="asis", fig.align="left", out.width="800px", out.height="300px"}

fig1

#write.csv(dtafig1, "X_COVID19_US_AgeComposition/data_figure1.csv")

```

## Figure 2. Trend of age-distribution in weekly COVID-19 cases, hospitalizations, and deaths by race/ethnicity: Hispanic, non-Hispanic black, vs. Non-Hispanic white 
```{r fig2, results="asis", fig.align="left", out.width="800px", out.height="800px"}
subplot(
        fig2case %>% 
            layout(yaxis=list(title = "% of weekly cases", 
                              font=list(size=8)),
                   showlegend = FALSE), 
        fig2hosp %>% 
            layout(yaxis=list(title = "% of weekly hospitalizations",
                              font=list(size=8)),
                   showlegend = FALSE), 
        fig2death %>% 
            layout(yaxis=list(title = "% of weekly deaths", 
                              font=list(size=8))
                   ), 
        nrows=3, shareX = TRUE , shareY=FALSE, titleY = TRUE)

#write.csv(dtafig2case, "X_COVID19_US_AgeComposition/data_figure2case.csv")
#write.csv(dtafig2hosp, "X_COVID19_US_AgeComposition/data_figure2hospitalization.csv")
#write.csv(dtafig2death, "X_COVID19_US_AgeComposition/data_figure2death.csv")
```

##__Figure 3. Trends of weekly case fatality ratio - see three options__ 

* Option 1: By race (Hispanic, Black, White - following figure 2 structure. same legends)
* Option 2: By age - among Hispanic, Black, White (breaking from the figure 2 structure, which is okay since this is not distribution anymore)
* Option 3: By age - among all races 

__Decision points/questions__:
1. currently all options starts from early April, to avoid idiosyncratic values in the beginning of the year. But there are still some (especially in option 3). __Should we start from May?__ 
2. currently 0-9 and 10-19 excluded (in options 2 & 3). __Should we drop more younger age-group, and focus on older age-groups?__ 
3. which option? 

##__Figure 3. Trends of weekly case fatality ratio (option 1)__ 
```{r fig3option1, results="asis", fig.align="left", out.width="800px", out.height="350px"}

dtasortedbyage %>% 
    select(racegroup, weekdate, age, cfr)%>%
    
    filter(racegroup=="Hispanic" | racegroup=="NH Black" | racegroup=="NH White")%>%
    spread(age, cfr, fill = NA, convert = FALSE) %>%
    group_by(racegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 1, shareX = TRUE, shareY = TRUE, margin=0.05)%>%
    layout(
        showlegend = FALSE,
        yaxis = list(title = "Three-week lagged CFR (per 100), log scale",
                     titlefont=list(size=12),
                     type = "log", range = c(-2, 2) )
        )  

```

##__Figure 3. Trends of weekly case fatality ratio (option 2)__ 

log scale vs. linear scale

```{r fig3option2, results="asis", fig.align="left", out.width="800px", out.height="400px"}

dtasortedbyrace %>% 
    select(racegroup, weekdate, agegroup, cfr)%>%
    
    mutate(racegroup=gsub("\\s", "", racegroup))%>%
    mutate(agegroup=gsub(" Years", "", agegroup))%>%
    mutate(agegroup=gsub("\\s", "", agegroup))%>%
    filter(agegroup!="0-9" & agegroup!="10-19" )%>%
    filter(racegroup=="cfrHispanic" | racegroup=="cfrNHBlack" | racegroup=="cfrNHWhite")%>%
    spread(racegroup, cfr, fill = NA, convert = FALSE) %>%
    group_by(agegroup) %>%
    do(p = panelbyrace(.)) %>%
    subplot(nrows = 1, shareX = TRUE, shareY = TRUE)%>%
    layout(
        
        yaxis = list(title = "Three-week lagged CFR (per 100), log scale",
                     titlefont=list(size=12),
                     type = "log", range = c(-2, 2) )
        )  
```

```{r fig3option2_linear, results="asis", fig.align="left", out.width="800px", out.height="400px"}

dtasortedbyrace %>% 
    select(racegroup, weekdate, agegroup, cfr)%>%
    
    mutate(racegroup=gsub("\\s", "", racegroup))%>%
    mutate(agegroup=gsub(" Years", "", agegroup))%>%
    mutate(agegroup=gsub("\\s", "", agegroup))%>%
    filter(agegroup!="0-9" & agegroup!="10-19" )%>%
    filter(racegroup=="cfrHispanic" | racegroup=="cfrNHBlack" | racegroup=="cfrNHWhite")%>%
    spread(racegroup, cfr, fill = NA, convert = FALSE) %>%
    group_by(agegroup) %>%
    do(p = panelbyrace_linear(.)) %>%
    subplot(nrows = 1, shareX = TRUE, shareY = FALSE)%>%
    layout(
        
        yaxis = list(title = "Three-week lagged CFR (per 100)",
                     titlefont=list(size=12) )
        )  
```

##__Figure 3. Trends of weekly case fatality ratio (option 3)__ 

log scale vs. linear scale

```{r fig3option3, results="asis", fig.align="left", out.width="800px", out.height="400px"}

dtasortedbyrace %>% 
    select(racegroup, weekdate, agegroup, cfr)%>%
    
    mutate(racegroup=gsub("\\s", "", racegroup))%>%
    mutate(agegroup=gsub(" Years", "", agegroup))%>%
    mutate(agegroup=gsub("\\s", "", agegroup))%>%
    filter(agegroup!="0-9" & agegroup!="10-19" )%>%
    spread(racegroup, cfr, fill = NA, convert = FALSE) %>%
    group_by(agegroup) %>%
    do(p = panelbyraceall(.)) %>%
    subplot(nrows = 1, shareX = TRUE, shareY = TRUE)%>%
    layout(
        
        yaxis = list(title = "Three-week lagged CFR (per 100), log scale",
                     titlefont=list(size=12),
                     type = "log", range = c(-2, 2) )
        )  
```

```{r fig3option3_linear, results="asis", fig.align="left", out.width="800px", out.height="400px"}

dtasortedbyrace %>% 
    select(racegroup, weekdate, agegroup, cfr)%>%
    
    mutate(racegroup=gsub("\\s", "", racegroup))%>%
    mutate(agegroup=gsub(" Years", "", agegroup))%>%
    mutate(agegroup=gsub("\\s", "", agegroup))%>%
    filter(agegroup!="0-9" & agegroup!="10-19" )%>%
    spread(racegroup, cfr, fill = NA, convert = FALSE) %>%
    group_by(agegroup) %>%
    do(p = panelbyraceall_linear(.)) %>%
    subplot(nrows = 1, shareX = TRUE, shareY = FALSE)%>%
    layout(
        
        yaxis = list(title = "Three-week lagged CFR (per 100)",
                     titlefont=list(size=12))
        )  
```


